Cover Letter: Cudele

### 2017-12-29 paper/sections/5_implementation.tex
paper: #41 and #34
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

In CephFS, the metadata store is a data structure that represents the file
system namespace. This data structure is stored in two places: in memory ({\it
i.e.} in the collective memory of the metadata server cluster) and as objects
in the object store. In the object store, directories and their inodes are
stored together in objects to improve the performance of scans.  The metadata
store data structure is structured as a tree of directory fragments making it
easier to read and traverse.

% how i used it
\textbf{Cudele}: the RPCs mechanism uses the metadata store to
service requests. Using code designed for recovery, Nonvolatile Apply and
Volatile Apply replay updates onto the metadata store in memory and in the
object store, respectively.  When the clients are ready to merge their updates
back into the global namespace, they pass a binary file of  metadata updates to
the metadata server.

@@

\subsection{Journal}

The journal is the second way that CephFS represents the
file system namespace; it is a log of metadata updates that can materialize the
namespace when the updates are replayed onto the metadata store. The journal is
a ``pile system"; writes are fast but reads are slow because state must be
reconstructed.  Specifically, reads are slow because there is more state to
read, it is unorganized, and many of the updates may be redundant.

\textbf{Cudele}: the journal format is used by Stream, Append Client Journal,
Local Persist, and Global Persist.  Stream is the default
implementation for achieving global durability in CephFS but the rest of the
mechanisms are implemented by writing with the journal format.  By writing with
the same format, the metadata servers can read and use the recovery code to
materialize the updates from a client's decoupled namespace ({\it i.e.} merge).
These implementations required no changes to CephFS because the metadata
servers know how to read the events the library is writing.  By re-using the
journal subsystem to implement the namespace decoupling, Cudele leverages the
write/read optimized data structures, the formats for persisting events
(similar to TableFS's SSTables~\cite{ren:atc2013-tablefs}), and the functions
for replaying events onto the internal namespace data structures.

@@

\subsection{Journal Tool}

The journal tool is used for disaster recovery and lets administrators view and
modify the journal. It can read the journal, export the journal as a file,
erase events, and apply updates to the metadata store.  To apply journal
updates to the metadata store, the journal tool reads the journal from object
store objects and replays the updates on the metadata store in the object
store.

\textbf{Cudele}: the external library the clients link into is based
on the journal tool. It already had functions for importing, exporting, and
modifying the updates in the journal so we re-purposed that code to implement
Append Client Journal, Volatile Apply, and Nonvolatile Apply.

@@

\subsection{Inode Cache}

In CephFS, the inode cache reduces the number of RPCs between clients and
metadata servers. Without contention clients can resolve metadata reads locally
thus reducing the number of operations ({\it e.g.}, \texttt{lookup()}s).  For
example, if a client or metadata server is not caching the directory inode, all
creates within that directory will result in a lookup and a create request. If
the directory inode is cached then only the create needs to be sent.  The size
of the inode cache is configurable so as not to saturate the memory on the
metadata server -- inodes in CephFS are about 1400
bytes\footnote{http://docs.ceph.com/docs/jewel/dev/mds\_internals/data-structures/}.
The inode cache has code for manipulating inode numbers, such as pre-allocating
inodes to clients.

\textbf{Cudele}: Nonvolatile Apply uses the internal inode cache
code to allocate inodes to clients that decouple parts of the namespace and to
skip inodes used by the client at merge time.

@@

\subsection{Large Inodes}

In CephFS, inodes already store policies, like how the file is striped
across the object store or for managing subtrees for load balancing. These
policies adhere to logical partitionings of metadata or data, like Ceph pools and file
system namespace subtrees. To implement this, the namespace data structure has
the ability to recursively apply policies to subtrees and to isolate subtrees
from each other.

\textbf{Cudele}: uses the large inodes to store consistency and
durability policies in the directory inode. This approach uses the File Type
interface from the Malacology programmable store
system~\cite{sevilla:eurosys17} and it tells clients how to access the
underlying metadata. The underlying implementation stores executable code in
the inode that calls the different Cudele mechanisms. Of course, there are many
security and access control aspects of this approach but that is beyond the
scope of this paper.

>>>>

In CephFS, the metadata store is a data structure that represents the file
system namespace. This data structure is stored in two places: in memory ({\it
i.e.} in the collective memory of the metadata server cluster) and as objects
in the object store. In the object store, directories and their inodes are
stored together in objects to improve the performance of scans.  The metadata
store data structure is structured as a tree of directory fragments making it
easier to read and traverse. \textbf{In Cudele} the RPCs mechanism uses the
metadata store to service requests. Using code designed for recovery,
Nonvolatile Apply and Volatile Apply replay updates onto the metadata store in
memory and in the object store, respectively.  When the clients are ready to
merge their updates back into the global namespace, they pass a binary file of
metadata updates to the metadata server.

@@

\subsection{Journal Format and Journal Tool}

The journal is the second way that CephFS represents the file system namespace;
it is a log of metadata updates that can materialize the namespace when the
updates are replayed onto the metadata store. The journal is a ``pile system";
writes are fast but reads are slow because state must be reconstructed.
Specifically, reads are slow because there is more state to read, it is
unorganized, and many of the updates may be redundant.  \textbf{In Cudele} the
journal format is used by Stream, Append Client Journal, Local Persist, and
Global Persist.  Stream is the default implementation for achieving global
durability in CephFS but the rest of the mechanisms are implemented by writing
with the journal format.  By writing with the same format, the metadata servers
can read and use the recovery code to materialize the updates from a client's
decoupled namespace ({\it i.e.} merge).  These implementations required no
changes to CephFS because the metadata servers know how to read the events the
library is writing.  By re-using the journal subsystem to implement the
namespace decoupling, Cudele leverages the write/read optimized data
structures, the formats for persisting events (similar to TableFS's
SSTables~\cite{ren:atc2013-tablefs}), and the functions for replaying events
onto the internal namespace data structures.

@@

The journal tool is used for disaster recovery and lets administrators view and
modify the journal. It can read the journal, export the journal as a file,
erase events, and apply updates to the metadata store.  To apply journal
updates to the metadata store, the journal tool reads the journal from object
store objects and replays the updates on the metadata store in the object
store.  \textbf{In Cudele} the external library the clients link into is based
on the journal tool. It already had functions for importing, exporting, and
modifying the updates in the journal so we re-purposed that code to implement
Append Client Journal, Volatile Apply, and Nonvolatile Apply.

@@

\subsection{Inode Cache and Large Inodes}

In CephFS, the inode cache reduces the number of RPCs between clients and
metadata servers. Without contention clients can resolve metadata reads locally
thus reducing the number of operations ({\it e.g.}, \texttt{lookup()}s).  For
example, if a client or metadata server is not caching the directory inode, all
creates within that directory will result in a lookup and a create request. If
the directory inode is cached then only the create needs to be sent.  The size
of the inode cache is configurable so as not to saturate the memory on the
metadata server -- inodes in CephFS are about 1400
bytes~\cite{docs:cephinternalds}.  The inode cache has code for manipulating
inode numbers, such as pre-allocating inodes to clients. \textbf{In
Cudele},Nonvolatile Apply uses the internal inode cache code to allocate inodes
to clients that decouple parts of the namespace and to skip inodes used by the
client at merge time.

@@

In CephFS, inodes already store policies, like how the file is striped across
the object store or for managing subtrees for load balancing. These policies
adhere to logical partitionings of metadata or data, like Ceph pools and file
system namespace subtrees. To implement this, the namespace data structure has
the ability to recursively apply policies to subtrees and to isolate subtrees
from each other. \textbf{In Cudele} the large inodes store consistency and
durability policies in the directory inode. This approach uses the File Type
interface from the Malacology programmable store
system~\cite{sevilla:eurosys17} and it tells clients how to access the
underlying metadata. The underlying implementation stores executable code in
the inode that calls the different Cudele mechanisms. Of course, there are many
security and access control aspects of this approach but that is beyond the
scope of this paper.

### 2017-12-29 paper/sections/6_results.tex
paper: #41 and #34
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\newcomment{\indent We run experiments on the same cluster and graph the
standard deviations for three runs (sometimes error bars are too small to see).
We also normalize results to control for hardware differences and to make our
results more generally applicable. We use a CloudLab cluster of 34 nodes
connected with 10Gbit ethernet, each with 16 2.4 GHz CPUs, 128GB RAM, and 400GB
SSDs.  Each node uses Ubuntu 14.04 and we develop on Ceph's Jewel release,
version 10.2.1, which was released in May 2016. All daemons run as a single
process (the default setting for Ceph) and the Ceph
components\footnote{http://docs.ceph.com/docs/jewel/rados/} are: 1 monitor, 3
object storage daemons, 1 metadata server, and 20 clients.} We scope the
evaluation to one metadata server and scale the number of parallel clients
\newcomment{each doing 100K operations}\footnote{This is the maximum
recommended size of a directory in CephFS; preliminary experiments with larger
directory sizes show memory problems.} \newcomment{We scale to 20 clients
because, as shown in Section~\S\ref{sec:posix-overheads}, 20 clients is enough
to saturate the resources of a single metadata server.} This type of analysis
shows the capacity and performance of a metadata server with superior metadata
protocols, which should be used to inform \oldcomment{load
balancing}\newcomment{metadata distribution} across a \oldcomment{metadata}
cluster. Load balancing across a cluster of metadata servers with partitioning
and replication can be explored with something like
Mantle~\cite{sevilla:sc15-mantle}.

\newcomment{\indent To make our results reproducible, this paper adheres to The
Popper Convention~\cite{jimenez_popper_2016}, so experiments presented here are
available in the repository for this article.  Experiments can be examined in
more detail, or even re-run, by visiting the \texttt{[source]} link next to
each figure. That link points to a Jupyter notebook that shows the analysis and
source code for that graph, which points to an experiment and its artifacts.
The source code is available on the Cudele
branch\footnote{\newcomment{https://github.com/michaelsevilla/ceph/tree/cudele}}
of our Ceph fork.} \addressesissue{1}

@@

\subsection{Microbenchmarks}
\label{sec:microbenchmarks}
\begin{figure}[tb]
\centering
\includegraphics[width=1.0\linewidth]{graphs/composable-mechanisms.png}
\caption{
[\href{https://github.com/michaelsevilla/cudele-popper/blob/revision/experiments/cudele-mechanisms/visualize/viz.ipynb}{source}]
\newcomment{Overhead of processing 100K create events for each mechanism in
Table~\ref{table:mechanisms}. Results are normalized to the runtime of writing
events to client memory. The bars on the far right show the overhead of
building consistency/durability semantics of real world systems.}
\oldcomment{Performance of each mechanism (left) and building the consistency/durability
semantics of real-world systems (right) for 100K files creates from a single
client. Results are normalized to the runtime of writing events to the client's
in-memory journal.}\label{fig:composable-mechanisms}}
\end{figure}

@@

% persist vs. save: one disk vs. many
{\it Parallelism of the Object Store:} \newcomment{ Stream\footnote{This
measurement is an approximation of the overhead and is calculated by
subtracting the runtime of the job with the journal turned off from the runtime
with the journal turned on.} has the highest slowdown at \(2.4\times\)
because the overhead of maintaining and streaming the journal is incurred by
the metadata server}.  Comparing Local and Global Persist demonstrates the
bandwidth advantages of storing the journal in a distributed object store. The
Global Persist performance is \newcomment{only \(0.2\times\) slower than Local
Persist because Global Persist}\oldcomment{\(1.5\times\) faster because the
object store} is leveraging the collective bandwidth of the disks in the
cluster. This benefit comes from the object store itself but should be
acknowledged when making decisions for the application; the bandwidth of the
object store can help mitigate the overheads of globally persisting metadata
updates. The storage per journal update is about 2.5KB. So the storage
footprint scales linearly with the number of metadata creates and suggests that
updates for a million updates in a single journal would be
\newcomment{2.38GB}\oldcomment{2.5KB \(*\) 1 million files \(=\) 2.38GB.
Transfer times for payloads of this size in most HPC/data center networks are
reasonable.}

@@

\newcomment{ ``Decoupled: merge + create" outperforms ``RPCs" by \(3.37\times\)
because ``decoupled: merge + create" uses a relaxed form of consistency and
leverages bulk updates just like DeltaFS~\cite{zheng:pdsw2015-deltafs}.
Decoupled namespaces (1)} place no restrictions on the validity of
metadata inserted into the journal ({\it e.g.}, never checking for the
existence of files before creating files), \newcomment{(2)} avoid touching
poorly scaling data structures\newcomment{, and (3) allow clients to batch
events into bulk updates}. \oldcomment{We can scale linearly if we weaken our
consistency by never checking for the existence of files before creating files
({\it i.e.} only append \texttt{open()} requests to the journal).  When the
updates are merged by the metadata server into the in-memory metadata store
they never scan growing data structures.}  Had we implemented the
\oldcomment{Append Client Journal mechanism}\newcomment{client to send
updates one at a time and} to include \texttt{lookup()} commands before
\texttt{open()} requests, we would have seen \newcomment{performance closer to
the ``RPC" subtree}\oldcomment{the poor scaling that we see with the RPCs
mechanism}.

\newcomment{One final point about the ``decoupled: merge + create" curve is
that the results in the graph are pessimistic. They model a scenario in which
all client journals arrive at the same time.} So for the 20 clients data point,
we are measuring the operations per second for 20 client journals that land on
the metadata server at the same time. \newcomment{Had we added infrastructure
to overlay journal arrivals or time client sync intervales, we could have
scaled more closely to ``decoupled: create".}\oldcomment{The ``create + merge"
bar adds the time it takes for clients to generate the journal of events in
parallel. ALSO DISCUSS 3.7X!!}

@@

\subsection{Discussion and Future Work}

\newcomment{We have shown that our API and framework is an effective
abstraction for letting \oldcomment{users}\newcomment{administrators} custom
fit subtrees to applications.  It is more scalable than the current practice of
mounting different storage systems in a global namespace because there is no
need to provision dedicated storage clusters to applications or move data
between these systems.  For example, the results of a Hadoop job do not need to
be migrated into a Ceph file system (CephFS) for other processing; instead the
\oldcomment{user}\newcomment{administrator} can change the semantics of the HDFS
subtree into a CephFS subtree.  This may cause metadata/data movement to make
things strongly consistent again but this is superior to moving all data across
file system boundaries. Our prototype enables studies that adjust these
semantics over {\it time and space}, where subtrees can change their semantics
and migrate around the cluster without ever moving the data they reference.}

Cudele and the experiments we present here prompt many avenues for future work.
First is quantifying the benefits of dynamically changing the semantics of a
subtree from stronger to weaker guarantees (or vice versa), as described in the
introduction. Second is embeddable policies, where child subtrees have
specialized features but still maintain the guarantees  of their parent
subtrees. We already show how applications can compose guarantees and control
their performance, knobs that we did not have before, but embeddable policies
are a way to compose the policies themselves.  For example, a RAMDisk subtree
is POSIX IO-compliant but relaxes durability constraints, so it can reside
under a POSIX IO subtree that is strongly consistent. These embeddable policies
should be controlled and enforced by Cudele.  Finally, performance prediction
benefits from our cost quantification of each mechanism. Applications can use
Cudele to microbenchmark their components and software, similar to what we did
in Section~\S\ref{sec:microbenchmarks}. Using those results, they can predict
how much slower their system will be if they adopt stronger consistency or
durability.  This is a form of co-design that takes a ``dirty-slate" approach
but building just the guarantees the application needs from existing
implementations.  This can also be a verification tool where performance that
varies wildly from the predicted performance can be a red flag that something
is wrong or that the bottleneck is not in the consistency or durability plane.

>>>>

\newcomment{\indent We graph standard deviations for three runs (sometimes
error bars are too small to see) and normalize results to make our results more
generally applicable to different hardware. We use a CloudLab cluster of 34
nodes connected with 10Gbit ethernet, each with 16 2.4 GHz CPUs, 128GB RAM, and
400GB SSDs.  Each node uses Ubuntu 14.04 and we develop on Ceph's Jewel
release, version 10.2.1, which was released in May 2016. All daemons run as a
single process (and the Ceph components~\cite{docs:cephrados} are: 1 monitor, 3
object storage daemons, 1 metadata server, and 20 clients.} We scope the
evaluation to one metadata server and scale the number of parallel clients
\newcomment{each doing 100K operations because 100K is the maximum recommended
size of a directory in CephFS. We scale to 20 clients because, as shown in
Section~\S\ref{sec:posix-overheads}, 20 clients is enough to saturate the
resources of a single metadata server.} This type of analysis shows the
capacity and performance of a metadata server with superior metadata protocols,
which should be used to inform \oldcomment{load balancing}\newcomment{metadata
distribution} across a \oldcomment{metadata} cluster. Load balancing across a
cluster of metadata servers with partitioning and replication can be explored
with something like Mantle~\cite{sevilla:sc15-mantle}.  \newcomment{To make our
results reproducible, this paper adheres to The Popper
Convention~\cite{jimenez_popper_2016}, so experiments can be examined in more
detail, or even re-run, by visiting the \texttt{[source]} link next to each
figure.  The source code is available on the Cudele branch~\cite{cudele-src} of
our Ceph fork.}

@@

\subsection{Microbenchmarks}
\label{sec:microbenchmarks}
\begin{figure}[tb]
\centering
\includegraphics[width=1.0\linewidth]{graphs/composable-mechanisms.png}
\caption{
[\href{https://github.com/michaelsevilla/cudele-popper/blob/revision/experiments/cudele-mechanisms/visualize/viz.ipynb}{source}]
\newcomment{Overhead of processing 100K create events for each mechanism in
\oldcomment{Table~\ref{table:mechanisms}}\newcomment{Figure~\ref{fig:decouple}},
normalized to the runtime of writing events to client memory. The far right
graph shows the overhead of building semantics of real world systems.}
\oldcomment{Performance of each mechanism (left) and building the
consistency/durability semantics of real-world systems (right) for 100K files
creates from a single client. Results are normalized to the runtime of writing
events to the client's in-memory journal.}\label{fig:composable-mechanisms}}
\end{figure}

@@

% persist vs. save: one disk vs. many
{\it Parallelism of the Object Store:}
\newcomment{Stream, which is an approximation (journal on minus journal off) has
the highest slowdown at \(2.4\times\) because the overhead of maintaining and
streaming the journal is incurred by the metadata server}.  Comparing Local and
Global Persist demonstrates the bandwidth advantages of storing the journal in
a distributed object store. The Global Persist performance is \newcomment{only
\(0.2\times\) slower than Local Persist because Global
Persist}\oldcomment{\(1.5\times\) faster because the object store} is
leveraging the collective bandwidth of the disks in the cluster. This benefit
comes from the object store itself but should be acknowledged when making
decisions for the application; the bandwidth of the object store can help
mitigate the overheads of globally persisting metadata updates. The storage per
journal update is about 2.5KB. So the storage footprint scales linearly with
the number of metadata creates and suggests that updates for a million updates
in a single journal would be \newcomment{2.38GB}\oldcomment{2.5KB \(*\) 1
million files \(=\) 2.38GB.  Transfer times for payloads of this size in most
HPC/data center networks are reasonable.}

@@

\newcomment{ ``Decoupled: merge + create" outperforms ``RPCs" by \(3.37\times\)
because ``decoupled: merge + create" uses a relaxed form of consistency and
leverages bulk updates just like DeltaFS~\cite{zheng:pdsw2015-deltafs}.
Decoupled namespaces (1)} place no restrictions on the validity of metadata
inserted into the journal ({\it e.g.}, never checking for the existence of
files before creating files), \newcomment{(2)} avoid touching poorly scaling
data structures\newcomment{, and (3) allow clients to batch events into bulk
updates}. \oldcomment{We can scale linearly if we weaken our consistency by
never checking for the existence of files before creating files ({\it i.e.}
only append \texttt{open()} requests to the journal).  When the updates are
merged by the metadata server into the in-memory metadata store they never scan
growing data structures.}  Had we implemented the \oldcomment{Append Client
Journal mechanism}\newcomment{client to send updates one at a time and} to
include \texttt{lookup()} commands before \texttt{open()} requests, we would
have seen \newcomment{performance closer to the ``RPC" subtree}\oldcomment{the
poor scaling that we see with the RPCs mechanism}. \newcomment{The ``decoupled:
merge + create" curve is also pessimistic because it models a scenario in which
all client journals arrive at the same time.} So for the 20 clients data point,
we are measuring the operations per second for 20 client journals that land on
the metadata server at the same time. \newcomment{Had we added infrastructure
to overlay journal arrivals or time client sync intervales, we could have
scaled more closely to ``decoupled: create".}\oldcomment{The ``create + merge"
bar adds the time it takes for clients to generate the journal of events in
parallel. ALSO DISCUSS 3.7X!!}

@@

%\subsection{Discussion and Future Work}
%
%\newcomment{We have shown that our API and framework is an effective
%abstraction for letting \oldcomment{users}\newcomment{administrators} custom
%fit subtrees to applications.

### 2017-12-29 paper/sections/7_conclusion.tex
paper: #41 and #34
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\section{Conclusion}

Relaxing consistency/durability semantics in the file system is a double-edged
sword. While it performs and scales better, it alienates applications that rely
on strong consistency and durability. Mounting other systems to the global
namespace is convenient but wastes resources and incurs data movement.  Cudele
lets users assign consistency/durability guarantees to subtrees in the global
namespace, which can be custom fit to the application. Using Cudele, we show
how applications can co-exist and perform well in a global namespace and lay
the groundwork for new systems that dynamically change consistency/durability
guarantees to avoid provisioning dedicated storage clusters and moving large
amounts of data.

>>>>

\section{Conclusion and Future Work}

Relaxing consistency/durability semantics in file systems is a double-edged
sword. While it performs and scales better, it alienates applications that rely
on strong consistency and durability.  Cudele lets
\oldcomment{users}\newcomment{administrators} assign consistency/durability
guarantees to subtrees in the global namespace, which lets administrators
custom fit semantics to application. We show how applications can co-exist and
perform well in a global namespace and our prototype enables studies that
adjust these semantics over {\it time and space}, where subtrees can change
without ever moving the data they reference.

%This approach is superior to mounting different storage systems in a global
%namespace because there is no need to provision dedicated storage clusters to
%applications or move data between these systems.  
Cudele prompts many avenues for future work.  First is dynamically changing
semantics of a subtree from stronger to weaker guarantees (or vice versa).  For
example, the results of a Hadoop job do not need to be migrated into CephFS for
other processing; instead the \oldcomment{user}\newcomment{administrator} can
change the semantics of the HDFS subtree into a CephFS subtree, which may cause
metadata/data movement to make ensure strong consistency. Second is embeddable
policies, where child subtrees have specialized features but still maintain
guarantees  of their parent subtrees.  For example, a RAMDisk subtree is POSIX
IO-compliant but relaxes durability constraints, so it can reside under a POSIX
IO subtree that is strongly consistent.

%Finally, our prototype enables
%performance prediction because it helps administrators quantify the costs of
%each mechanism (similar to Section~\S\ref{sec:microbenchmarks}).

%Applications can use Cudele to microbenchmark their components and software,
%similar to what we did in

%Using those
%results, they can predict how much slower their system will be if they adopt
%stronger consistency or durability.  This is a form of co-design that takes a
%``dirty-slate" approach but building just the guarantees the application needs
%from existing implementations.  
%This can also be a verification tool where performance that varies wildly from
%the predicted performance can be a red flag that something is wrong or that
%the bottleneck is not in the consistency or durability plane.
